\section*{Acknowledgements}
I would like to express my gratitude to my supervisor, \textbf{Prof. Stefano Pagani}, for his support and guidance throughout this project. I definitely need to thank my co-advisor, \textbf{Dr. Stéphane Cotin}, for his insights and expertise, but also for the opportunity to work with the Mimesis team at INRIA in Strasbourg. While at INRIA, I had the chance to work with many talented people, which helped me shape my work and my ideas. Along with Stéphane, I would like to thank \textbf{Dr. Nicola Zotto} and \textbf{Dr. François Lecomte} for their valuable feedback and discussions. Everyone in the Mimesis team has been very welcoming and helpful, making my time there an unforgettable experience. 

This work would not have been possible without the support of my family and friends. In particular, I would like to thank \textbf{ZioBar} with whom I shared many beautiful memories, the group \textbf{TuttoAttaccatoTest} for being the most important thing I discovered during my Master's degree, and my friends \textbf{Jacopo} and \textbf{Riccardo} for countless nights in Piazza Leonardo da Vinci.

I want to express my deepest gratitude to my parents, for their constant support and encouragement throughout my academic journey. They definitely helped me become the person I am today, and I am forever grateful for their love and sacrifices. Also to my siblings, \textbf{Martina}, \textbf{Matteo}, and \textbf{Francesco}, for being always there for me, I am very lucky to have you in my life.

Finally, a special thanks to \textbf{Ilaria}. While this journey has come to an end, I am excited to start a lot of new ones with you. Words will never be enough to express how grateful I am to have you by my side. Thanks for everything.