\section{Lesson 12/10/2022}
Typewriter sequence che però aveva iniziato la lezione scorsa

\begin{remark}
    \( f_p \nrightarrow 0 \) a.e. on \(\left[ 0, 1\right]\).  
    But consider \( \{ f_{p(n,1)}: n \in \mathbb{N} \} \). 
    This is a subsequence and, by definition \( f_{p(n, 1)}(x) = \chi_{n, 1}(x)= \chi_{\left[0, \frac{0}{n} \right]}(x) \). 
    For this subsequence, we have \( f_{p(n,1)}(x) \rightarrow 0 \) as \( n \to\infty \; \forall x \in (0, 1] \), then a.e. on \(\left[0, 1\right]\)
\end{remark}
This is not random!

\begin{proposition}
    If \(\mu(X) < \infty \) and \(f_n \rightarrow f \) in measure, then \(\exists\) a subsequence \(\{f_{n_k} \}\) s.t. \(f_{n_k} \to f \) a.e. on \(X\).
\end{proposition}
Now we analize the relation between convergence in \(L^1(X)\) and the other convergences.

\begin{theorem}
    \( \{f_n\} \subset L^1(X), f \in L^1(X) \). If \(f_n \rightarrow f \) in \(L^1(X)\) then \(f_n \rightarrow f \) in measure on \(X\)
\end{theorem}
\begin{proof}
    By contradiction. Suppose that \(f_n \nrightarrow f \) in measure on X: 
    \( \exists \bar{\alpha} > 0 \) s.t. 
    \[ \limsup_{n\to\infty} \mu(\{ |f_n-f| \geq \bar{\alpha} \}) > 0 \]
    \(\Rightarrow \exists \bar{\epsilon}\) and a subsequence \( \{ f_{n_k} \} \) s.t.
    \[ \mu(\{ |f_{n_k}-f| \geq \bar{\alpha} \}) > \bar{\epsilon} \]
    Consider then \(d_1(f_{n_k}, f)= \int_X |f_{n_k} - f| \, d\mu  \geq \int_{\left\{|f_{n_k}-f|\geq \bar{\alpha}  \right\}} 1 \, d\mu 
    = \bar{\alpha} \mu (\{|f_{n_k } - f| \geq \bar{\alpha}\}) > \bar{\alpha} \bar{\epsilon}\) 
    But, by assumption, \(d_1(f_n, f) \rightarrow 0\)
    \[ \Rightarrow d_1(f_{n_k}, f) \rightarrow 0 \] 
    contradiction.
\end{proof}

\begin{remark}
    the convergence in measure doesn't imply the convergence in \(L^1\). For example, consider 
    \( f_n (x) = n \chi_{\left[0, \frac{1}{n} \right]}(x) \)
    \( \mu \left( \left\{ |f_n| \geq \alpha \right\}\right) \to 0 \) for every \(\alpha\) \\
    On the other hand \( \int _{\left[0, 1\right]} n \chi_{\left[0, \frac{1}{n} \right]} \, d\lambda = \int_{\left[0, \frac{1}{n}\right]} n \, d\lambda = n \frac{1}{n} = 1\)
    \( f_n \nrightarrow 0\) in \(L^1\) 
\end{remark}


Convergence a.e. \(\nRightarrow\) convergence in \(L^1\): \\
use the same example above, \(f_n \rightarrow 0\) a.e. on \([0, 1] \nRightarrow f_n \rightarrow 0\) in \(L^1\)

Convergence in \(L^1\) \(\nRightarrow\) convergence a.e.
Consider the typewriter sequence: \( d_1(f_{p(n, k)}, 0) \to 0\) when \( p \to\infty\) \\
But we don't have a.e. convergence. However, recall the dominated convergence theorem: (DOM)
\[ f_n \rightarrow f \text{ a.e. }  + \exists \text{ of a dom function } \Rightarrow d(f_n, f)\rightarrow 0 \]
It is also possible to show a reverse DOM:
if \(f_n \to f \) in \(L^1(X)\), then \(\exists\) a subsequence \(\left\{f_{n_k}\right\}\) and \(w \in L^1(X)\) s.t. 
\begin{enumerate}
    \item \(f_{n_k} \rightarrow f\) a.e. on X
    \item \(\| f_{n_k} \| \leq w(x) \) for a.e. x \(\in X\)
\end{enumerate}

\subsection*{Derivatives of measures}
\(\left(X, \mathbb{M}, \mu \right)\) measure space. 
\(\phi : X \to \left[0, \infty \right]\) measurable. 
We learned that \(\nu: \mathbb{M} \to \left[0, \infty \right]\) by \[\nu(E)= \int_E \phi \, d\mu\] is a measure on \(X, \mathbb{M}\).

If the equation above holds, then we say that \(\phi \) is the Radon Nykodym derivative of \(\nu\) with respect to \(\mu\) and we write \[\phi = \frac{d\nu}{d\mu}\]
\begin{definition}
    \(\mu, \nu  \) measures on \(\left(x, \mathbb{M}\right)\). We say that \(\nu\) is absolutely continuous with respect to \(\mu\), \(\nu << \mu \) if 
    \[\mu(E) = 0 \Rightarrow \nu(E)=0\]
\end{definition}

\begin{lemma}
    There is a necessary condition:
    \[ \exists \frac{d \nu}{d \mu} \Rightarrow \nu << \mu \]
\end{lemma}

\begin{proof}
    \(\nu(E) = \int_E (\frac{d\nu}{d\mu}) \, d\mu = 0\) if \(\mu(E)=0\) by basic properties of \(\int\)
\end{proof}

\begin{theorem}[Radon Nykodim Theorem]
    \(\left(X, \mathbb{M}\right) \) measurable space, \(\mu, \nu\) measures. If \(\nu << \mu \) and moreover \(\mu \) is \(\sigma\) finite, then \(\phi : \to \left[0, \infty\right]\) measurable s.t.
    \(\phi = \frac{d \nu}{d \mu}\) namely \(\nu(E)= \int_E \phi \, d\mu \; \forall E \in \mathbb{M}\)
\end{theorem}

\begin{remark}
    if \(\mu\) is not sigma finite the theorem may fail.
    ..... 
\end{remark}

mi sto addormentando io ci sto provando anche
altre cose su radon-

\subsection*{Product Measure}
\( (X, \mathcal{M}, \mu), (Y, \mathcal{N}, \nu) \) measure spaces.
the goal is to define a measure space on \(X \times Y\)
\begin{definition}
    we call measurable rectangle in \(X \times Y\) a set of type \(A \times B\) where \(A \in \mathcal{M}, B \in \mathcal{N}\)
    \[  R = \{ A \times B \subset X\times Y ....\}\]
    We define the product \(\sigma\) algebra 
    ...
\end{definition}
\begin{definition}
    let \(E \subset X \times Y \) For \( \bar{x} \in X \) and \(\bar{y} \in Y \) we define
    \begin{enumerate}
        \item \( E_{\bar{x}} = \{ y \in Y: \left( \bar{x}, y \right) \in E \} \subset Y \)
        \item \( E_{\bar{y}} = \{ x \in X: \left( x, \bar{y} \right) \in Y \} \subset E \)
    \end{enumerate}
\end{definition}

\begin{proposition}
    \(\left( X, \mathcal{M} \right), \left( Y, \mathcal{N} \right)\) measurable spaces. 
    ....

\end{proposition}
...

\begin{theorem}
    If \(\left(X, \mathcal{M}, \mu \right)\) and \(\left(Y, \mathcal{N}, \nu \right)\) are \(\sigma\) finite spaces, then:
    \begin{enumerate}
        \item if \(\phi\) is \(\mathcal{M}\) measurable and \(\psi\) is \( \mathcal{N}\) meas
        \item we have that 
    \end{enumerate}

 \end{theorem}
using ....

\begin{theorem}
    iterated integrals for characteristic functions \\ 
    \(\mu \times \nu : \mathcal{M} \otimes \mathcal{N} \rightarrow \mathbb{R} \) defined by
    \[ \left(\mu \otimes \nu \right)(E) = \int_X \nu(E_x) \, d\mu = \int_Y \mu(E_y) \, d\nu\]
    is a measure, the product measure 
\end{theorem}

...
\begin{theorem}
    Let \(\lambda_n\) be the lebesgue measure in \(\mathbb{R}^n\). If \(n= K+m\), then \(\left(\mathbb{R}^n, \mathcal{L}(\mathbb{R}^n), \lambda_n \right)\) is the complection of \(\real^k \times \real^m .., \lambda_k \otimes \lambda_m\)
\end{theorem}
